---
title: "Visualizing NYC Traffic Accients Data"
output: html_document
---

Data on motor vehicle collisions in New York City has been made accesible by [NYC Open Data](<https://data.cityofnewyork.us/NYC-BigApps/NYPD-Motor-Vehicle-Collisions/h9gi-nx95)

``````{r, message=FALSE, warning=FALSE, results="hide"}
library(ggplot2);library(ggmap);library(dplyr);
library(reshape2);library(data.table);library(gridExtra);
```

### Munging the data 

```{r}
collisions<-read.csv('~/Desktop/NYPD_Motor_Vehicle_Collisions.csv')
collisions$DATE<-as.Date(collisions$DATE,"%m/%d/%Y")

columns<-which(colnames(collisions)%in%c("DATE","LATITUDE","LONGITUDE",
                                        "NUMBER.OF.PEDESTRIANS.INJURED","NUMBER.OF.PEDESTRIANS.KILLED",
                                        "NUMBER.OF.CYCLIST.INJURED","NUMBER.OF.CYCLIST.KILLED",
                                        "NUMBER.OF.MOTORIST.INJURED","NUMBER.OF.MOTORIST.KILLED"))

collisions<-collisions[,columns] 

colnames(collisions)<-c("Date","lat","lon",
                        "Pedesterian Injured","Pedesterian Killed",
                        "Cyclist Injured","Cyclist Killed",
                        "Motorist Injured","Motorist Killed")
head(collisions)

melt.collisions<-melt(collisions,id.vars=c('Date','lat','lon'))

melt.collisions$party<-""
melt.collisions$party[grep('Pedesterian',melt.collisions$variable)]<-"Pedestrian"
melt.collisions$party[grep('Cyclist',melt.collisions$variable)]<-"Cyclist"
melt.collisions$party[grep('Motorist',melt.collisions$variable)]<-"Motorist"
melt.collisions$outcome<-""
melt.collisions$outcome[grep('Injured',melt.collisions$variable)]<-"Injured"
melt.collisions$outcome[grep('Killed',melt.collisions$variable)]<-"Killed"

head(melt.collisions)

melt.collisions<-filter(melt.collisions,value>=1)
melt.collisions<-melt.collisions[complete.cases(melt.collisions),]
```

### Black and White Maps as 'Base Layer' of Visualization

``````{r, message=FALSE, warning=FALSE, results="hide",fig.height=8,fig.width=5.5}
nycmap_bw1 <- qmap("DeKalb Ave and Myrtle Ave, new york", zoom = 12, color = "bw", legend = "topleft")
grid.arrange(nycmap_bw1)

nycmap_bw2 <- qmap("1 E 161st St, Bronx, NY 10451", zoom = 12, color = "bw", legend = "topleft")
grid.arrange(nycmap_bw2)
```

### Overlay Layer to Visualize Fatal Accidents

```{r}
overlay <- stat_bin2d(aes(x = lon, y = lat, fill =party),size = .5, bins = 80, alpha = 2/3,data = filter(melt.collisions,outcome=="Killed"))
```

```{r,fig.height=8,fig.width=5.5}
map1<-nycmap_bw1+
  overlay+
  theme(legend.position="bottom")+
  scale_fill_hue(l=40)+
  ggtitle("Lower Manhattan/Queens/Brooklyn")+
  labs(fill='Type')

map2<-nycmap_bw2+
  overlay+
  theme(legend.position="bottom")+
  scale_fill_hue(l=40)+
  ggtitle("Upper Manhattan/Queens/Bronx")+
  labs(fill='Type')

```

### Finally, use gridExtra package to put two maps together

```{r,fig.height=10,fig.width=16}

grid.arrange(map1,map2,ncol=2,main=textGrob("Fatal Traffic Accident in NYC\nData from NYC Open Data   Visualization by @AKirayoglu", gp=gpar(cex=1.25), just="top"))

```

R markdown code is [here](https://github.com/alphankirayoglu/NYC_traffic_collisions/blob/master/Traffic_collisions.Rmd)
